---
title: "stats"
author: "Alex"
date: "1/2/2022"
# output: html_document
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# WikiPathways Stats
This R notebooks prepares figures to summarize WikiPathways activity. The output files are displayed on the website and used in publications and grant applications. Please edit in coordination with the WikiPathways development team.

* Data points are collected in _*data/*
* Plots are saved in *assets/img/*
 
```{r install, include=FALSE}
library(rWikiPathways) 
library(RColorBrewer)
library(scales)
library(plyr)
library(tidyverse)
```

## Collect Data

### Initialize Pathway Count Table
Collect history from old webservice, using getPathwayHistory on "Approved" 
pathways and checking oldest revision on pathways after WP3959 against a 
cutoff "oldest.date". 

**This was used to populate the data table; only run once.** See GitHub 
Collections to update the table using data sourced from GitHub repos.

```{r collect, eval=FALSE, include=FALSE}
wpid.all <- getPathwayIdsByCurationTag('Curation:AnalysisCollection')
oldest.date <- "201701"
oldest.wpid <- "WP3959" # manually confirmed
wpid.all.df <- plyr::ldply(wpid.all, function(p){
  if(p > oldest.wpid){
    wp.hist <- getPathwayHistory(p, oldest.date)
    if(length(wp.hist) > 0){
      t <- wp.hist[[1]]$timestamp
      if(!is.null(t)){
        data.frame(date=substr(t,0,6),wpid=p)
      }
      else
        data.frame(date=oldest.date,wpid=p)
    } else
      data.frame(date=oldest.date,wpid=p)
  } else
    data.frame(date=oldest.date,wpid=p)
})

## save raw data
write.table(wpid.all.df, "../_data/old/wpid_oldest_rev_2017-2022.csv", sep = ",", row.names = F)

## summarize counts
wpid.all.df.cnts <- wpid.all.df %>%
  dplyr::group_by(date) %>%
  dplyr:: summarise(pathways = n()) %>%
  dplyr::mutate(pathways = cumsum(pathways))

## save count data
write.table(wpid.all.df.cnts, "../_data/pathway_counts.csv", sep = ",", row.names = F)
```

### Initialize Monthly Activity Table
Collect monthly edit history from "User edits in month" from old server at 
*wpi/statistics/editCounts.txt*, manually simplified to *editCounts.csv*.

**This was used to populate the data table; only run once.** See GitHub 
Collections to update the table using data sourced from GitHub repos.

```{r edits, eval=FALSE, include=FALSE}
edits.user.df <- read.csv("../_data/old/editCounts.csv", stringsAsFactors = F)
edits.user.df$date <- as.character(strftime(strptime(edits.user.df$date,"%m/%d/%y"), "%Y%m"))

## save count data
write.table(edits.user.df, "../_data/edit_counts.csv", sep = ",", row.names = F)
```

### GitHub Collections

```{r github}
wpid.all.df.cnts <- read.csv("../_data/pathway_counts.csv", stringsAsFactors = F)
edits.user.df <- read.csv("../_data/edit_counts.csv", stringsAsFactors = F)

# TODO: count WP folders in _pathways/
# TODO: commits per time frame, see https://git-scm.com/docs/git-rev-list with
# parms --since and --until
```

## Plot data

Composite plot for main page: pathway count and number of edits per month.

First, let's combine our data frames and make a proper date column and factor by 
month
```{r prep}
combo.df <- edits.user.df %>%
  full_join(wpid.all.df.cnts, by="date")

combo.df$date <- strptime(paste0(combo.df$date,"01"), "%Y%m%d")
combo.df$month <- factor(format(combo.df$date, "%B"),
                                levels = month.name)
```

Next, let's plot a time series
```{r plot}

# RColorBrewer::display.brewer.all()
bcols <- RColorBrewer::brewer.pal(3,"Set1")

# scaling
ylim.prim <- c(0, max(combo.df$edits, na.rm = T)) # range for edits
ylim.sec <- c(min(combo.df$pathways, na.rm = T), max(combo.df$pathways, na.rm = T))    # range for pathways
b <- diff(ylim.prim)/diff(ylim.sec)
a <- b*(ylim.prim[1] - ylim.sec[1])

p <- ggplot(combo.df) +
  geom_bar(aes(x = as.Date(date),y=edits),stat="identity", fill=bcols[2]) +
  geom_line(data=combo.df[which(combo.df$month %in% c("February","December")),], # hand-picked dates due to patchy pathway counts
            aes(x = as.Date(date),y=a + pathways * b), 
            color = bcols[1], size = 2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",
               name = "",
               limits = c(as.Date(strptime("2018","%Y")),as.Date(strptime("2022","%Y")))) +
  scale_y_continuous(name="# Edits",  
                     sec.axis=sec_axis(~ (. - a)/b, 
                                       name="# Pathways")) +
  ggtitle("Active community and growing database") +
  xlab("") +
  theme(axis.text.y.left=element_text(colour=bcols[2]),
        axis.text.y.right=element_text(colour=bcols[1]),
        axis.ticks.y.left=element_line(colour=bcols[2]),
        axis.ticks.y.right=element_line(colour=bcols[1]))


p
```

```{r save}

ggsave("../assets/img/main_stats.png", plot = last_plot(), 
       width = 1100, height = 550, units = "px", dpi = 250)
```